A New Metric for the Analysis of Swarms that are modelled using potential fields

### 2018-08-05 paper.tex
MOre Updates

\title{A New Metric for the Analysis of Swarms}
\author{\uppercase{Neil Eliot}\authorrefmark{1},
\uppercase{David Kendall\authorrefmark{1}, and Michael Brockway\authorrefmark{1}}}
\address[1]{Northumbria University, Department of Computing and Information Sciences, 
Newcastle upon Tyne, NE1 8ST}

\markboth
{Eliot \headeretal: A New Metric for the Analysis of Swarms}
{Eliot \headeretal: A New Metric for the Analysis of Swarms}

@@

\begin{abstract}
  % An important criterion in the assessment of swarm coordination algorithms is
  % the extent to which they cause unnecessary movement, i.e. movement which
  % occurs as a side effect of the algorithm rather than from the fulfilment of
  % the intended function of the swarm. 
  A widely used metric for the analysis of swarm coordination algorithms is
  based upon the distance between agents. This paper examines this metric and
  introduces an entirely new metric based upon the field effects that createof the cohesion
  and repulsion vectors between agents. In comparing and contrasting the two
  metrics, we find that the \emph{cohesion/repulsion} metric offers a number of
  advantages over the distance metric. In particular, the
  \emph{cohesion/repulsion} metric allows the identification of the essential
  characteristic of a swarm as `expanding', `stable', or `contracting'. This is
  not possible with the distance metric.  Practical swarming applications where
  the new metric is useful include area filling and reconnaissance.
\end{abstract}

@@

Many coordination algorithms exist to control the movement of agents within a swarm. The primary goal of these algorithms is to maintain a stable swarm structure. The main focus of this paper is the use any coordination algorithm that are based on \textit{field effects} to produce a metric that allows the swarm to be analysed. To allow this paper to present the findings a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison. This includes the identification of a swarm's `state'. In the remainder of this paper Section~\ref{Section:Related} discusses related work, Section~\ref{Section:SwarmModelling} introduces a basic swarm model for evaluation. Section~\ref{Section:StabilityMagnitude} discusses the mathematical model that produces the swarming effect. Section~\ref{Section::Movement1} discusses how the model is implemented over time to produce movement. Section~\ref{Section:StabilityNullVector} discusses the effect of the model on stability. Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy. Section~\ref{Section:DistanceDynamics} discusses a distance based metric approach to swarm analysis. Section~\ref{Section:MagnitudeDynamics} introduces the new \textit{cohesion/repulsion-based} metric. Section~\ref{Section:ApplicationFloodFilling} compares the two metrics using a scenario of a swarm being used to fill an area. Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of the experiments and Section~\ref{Section:FutureWork} discuss additional work that needs to be carried out to exploit the additional information that the new metric provides.

@@

Many metrics exist for the analysis of swarms. These metrics can be categorised as \textit{Centroidal}, \textit{Inter-agent distance} and \textit{magnitude} based. The common denominator of the \textit{centroidal} and \text{inter-agent distance} techniques is that they \textit{measure} the agent distribution of a swarm based on distance. \textit{Magnitude} based analysis is based upon inter-agent interaction.

@@

\textit{Inter-Agent Distance} analysis is a basic mechanism for analysing a swarm it to analyse the separation of all of the agents with respect to each agents visible neighbours. This is different to using the centroid approach as the measure is based on agent interactions rather than a resultant structure. This technique is the most popular approach to analysing a swarms structure. Gazi and Passino use a distance based analysis in \cite{GP:04, GP:11} to identify optimum distribution of agents based upon different cohesion and repulsion applications. Wu et. al. discuss using several metrics to evaluate emergent behaviours in swarm based on a Viscsek model. These include calculating the average number of neighbours for each agent, grouping agents based on numbers of neighbours (to identify clusters). They also use the average distance between agents~\cite{WCW:11}.

@@

%% \section{Metric principles\label{section:MagnitudeDynamics}}
%% This section discusses the theory behind swarm movement and the effects that can be measured using both the metrics described above.
\section{Swarm modelling}\label{Section:SwarmModelling}
Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two field effects to implement the swarming characteristic. These effects are \emph{cohesion}, to draw agents closer, and \emph{repulsion} to prevent agents colliding. \emph{Field effects} have ranges around an agent and determine the effect other agents have upon it's movement~(Fig.~\ref{methods:FieldEffects}). It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$. When an agent ($b'$) moves into the \emph{neighbour field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion. When an agent $b'$ moves into the repulsion field of $b$ then $b$ also subjected to repulsion. When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e. repelled. When an agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion range $O_b$, it has it's repulsion vector applied and the agent will tend towards moving away from the obstacle.

@@

\section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in (\ref{eq:FlyToCentre1}) and (\ref{eq:Repulsion1}). Notice that the vectors are along the line of separation $bb'$. \\

@@

Equation~(\ref{eq:FlyToCentre1}) identifies the resultant cohesion effect of all the neighbours of $b$. Here, $nbr(b)$ defines the set of all the agents in the swarm that are a neighbour of $b$; a neighbour is any agent that fails within the cohesion field.

@@

Equation~(\ref{eq:Repulsion1}) identifies the resultant repulsion effect of all the neighbours of $b$ where $rep(.)$ returns a set of all agents that are within the repulsion field.

>>>>

\title{A New Metric for the Analysis of Swarms that are modelled using potential fields}
\author{\uppercase{Neil Eliot}\authorrefmark{1},
\uppercase{David Kendall\authorrefmark{1}, and Michael Brockway\authorrefmark{1}}}
\address[1]{Northumbria University, Department of Computing and Information Sciences, 
Newcastle upon Tyne, NE1 8ST}

\markboth
{Eliot \headeretal: A New Metric for the Analysis of Swarms that are modelled using potential fields}
{Eliot \headeretal: A New Metric for the Analysis of Swarms that are modelled using potential fields}

@@

\begin{abstract}
  % An important criterion in the assessment of swarm coordination algorithms is
  % the extent to which they cause unnecessary movement, i.e. movement which
  % occurs as a side effect of the algorithm rather than from the fulfilment of
  % the intended function of the swarm. 
  There are many metrics defined for the analysis of swarm coordination algorithms most of which are
  based upon the distances between agents or between agents and a fixed point. This paper examines a distance-based metric that measures a swarms overall structure using inter-agent distances and introduces an new metric designed to work in a similar way, at the global level, but is based upon the potential field effects that create the swarm's structure. The algorithms used to implement the swarming feature are based upon mathematically generated cohesion and repulsion vectors between an agent and its neighbours. In comparing and contrasting the two metrics, we find that the \emph{cohesion/repulsion} metric offers a number of advantages over the distance metric. In particular, the \emph{cohesion/repulsion} metric allows the identification of the essential characteristic of a swarm as `expanding', `stable', or `contracting'. This is not possible with a distance metric. Practical swarming applications where the new metric can be applied include area filling and reconnaissance.
\end{abstract}

@@

Many coordination algorithms exist to control the movement of agents within a swarm. The primary goal of these algorithms is to maintain a stable swarm structure. A stable structure is one where the agents in the swarm ``act'' as a single entity. The main focus of this paper is the use any coordination algorithms that are based on \textit{potential field effects} to produce a metric that allows the swarm to be analysed. To facilitate this analysis a basic swarm model is introduced and then the two metrics are applied to this model to present a comparison. This analysis includes the identification of a swarm's ``state''. The remainder of this paper is structured as follows Section~\ref{Section:Related} discusses different metrics that are currently used to analyse swarms, Section~\ref{Section:SwarmModelling} introduces a basic swarm model to produce an environment that can be analysed and evaluated using the two metrics. Section~\ref{Section:StabilityMagnitude} discusses the mathematical model that produces the swarming effect and how the potential fields are applied. Section~\ref{Section::Movement1} discusses how the model is implemented over time to produce movement. Section~\ref{Section:StabilityNullVector} discusses the effect of the model on swarm stability and the effects models have on swarm movements. Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy. Section~\ref{Section:DistanceDynamics} discusses a distance based metric approach to swarm analysis. Section~\ref{Section:MagnitudeDynamics} introduces the new \textit{cohesion/repulsion-based} metric and how it is calculated and applied. Section~\ref{Section:ApplicationFloodFilling} compares the two metrics using a scenario of a swarm being used to fill an area. Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of the experiments and Section~\ref{Section:FutureWork} discusses additional work that needs to be carried out to exploit the additional information that the new metric provides.

@@

Many metrics exist for the analysis of swarms. These metrics can be categorised as \textit{Centroidal}, \textit{Agent/Inter-agent distance} and \textit{magnitude} based. The common denominator of \textit{centroidal} and \text{Agent/inter-agent distance} metrics is that they \textit{measure} the agent distribution within a swarm based upon distances. \textit{Magnitude} based metrics analysis a swarm based upon inter-agent interaction.

@@

\textit{Inter-Agent Distance} analysis is a basic mechanism for analysing a swarm it to analyse the separation of all of the agents with respect to each agents visible neighbours. This is different to using the centroid approach as the measure is based on agent interactions rather than a resultant structure. This technique is the most popular approach to analysing a swarms structure. Gazi and Passino use a distance based analysis in \cite{GP:04, GP:11} to identify optimum distribution of agents based upon different cohesion and repulsion applications. Wu et. al. discuss using several metrics to evaluate emergent behaviours in swarm based on a Viscsek model. These include calculating the average number of neighbours for each agent, grouping agents based on numbers of neighbours (to identify clusters). They also use the average distance between agents~\cite{WCW:11}. Schneider and Wildermuth~\cite{SW:03} used a slightly different approach to measuring performance. They identify the difference between a simulated multi-robot environment and a physical implementation and identifying the difference in the paths of the robots based on distance variance.

@@

%% \section{Metric principles\label{section:MagnitudeDynamics}}
%% This section discusses the theory behind swarm movement and the effects that can be measured using both the metrics described above.
\section{Swarm modelling}\label{Section:SwarmModelling}
Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two field effects to implement the swarming characteristic. These effects are \emph{cohesion}, to draw agents closer, and \emph{repulsion} to prevent agents colliding. \emph{Potential fields} are ranges around an agent that determine the effect other agents have upon it's movement~(Fig.~\ref{methods:FieldEffects}). It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$. When an agent ($b'$) moves into the \emph{cohesion field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion. When an agent $b'$ moves into the \textit{repulsion field} of $b$ then $b$ is subjected to repulsion. When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e. repelled. When an agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion range $O_b$, it has it's repulsion vector applied and the agent will move away from the obstacle.

@@

\section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in Equation~\ref{eq:FlyToCentre1} and \ref{eq:Repulsion1}. Notice that the vectors are along the line of separation $bb'$. \\

@@

Equation~\ref{eq:FlyToCentre1} identifies the resultant cohesion effect of all the neighbours of $b$. Here, $nbr(b)$ defines the set of all the agents in the swarm that are a neighbour of $b$; a neighbour is any agent that fails within the cohesion field.

@@

Equation~\ref{eq:Repulsion1} identifies the resultant repulsion effect of all the neighbours of $b$ where $rep(.)$ returns a set of all agents that are within the repulsion field.
